Intentional release of native species undermines ecological stability

### 2022-10-01 rmd/manuscript.Rmd
Update manuscript.Rmd

Human demands for natural resources are ever-increasing, such that active interventions are critical to the sustainable management of fisheries, forestry, and wildlife [@laikreCompromisingGeneticDiversity2010]. Captive breeding is a form of the efforts to enhance wild populations of diverse plant and animal taxa [@snyderLimitationsCaptiveBreeding1996; @laikreCompromisingGeneticDiversity2010]. Although releases of captive-bred native species ("intentional release") entail ecological risks such as the accumulation of deleterious alleles [@arakiGeneticEffectsCaptive2007; @krkosekDecliningWildSalmon2007], this method is still pervasive in conservation [@snyderLimitationsCaptiveBreeding1996] and natural resource management [@laikreCompromisingGeneticDiversity2010]. In fisheries, for example, billions of hatchery individuals are released annually across the globe [@kitadaEconomicEcologicalGenetic2018]. The widespread use of captive breeding is perhaps because of the "myth" that the demographic benefits of released individuals may exceed the associated risks.

Current debates, however, overlook the fact that we have rarely assessed the community-wide impact of intentional release. Species are all embedded in the complex web of interacting organisms, and the stable coexistence of competing species through density-dependent feedback underpins the emergent stability of ecological communities [@tilmanEcologicalConsequencesChanges1999; @carrollNicheFitnessDifferences2011a; @loreauDoesFunctionalRedundancy2004]. Intentional release may disrupt the sensitive balance of species interactions because it introduces unnaturally high numbers of individuals into the wild [@careyNativeInvadersChallenges2012]. Hence, this form of species management may intervene in the ecological process that allows competing species to coexist, ultimately degrading the long-term community stability. Evidence for this hypothesis is lacking, however.

@@

We employed a multispecies Ricker model [@fowlerSpeciesDynamicsAlter2012] to simulate community dynamics with the selective intentional release of a constituent species (species 1). Specifically, the population density of species $i$ at time $t+1$, $N_{i,t+1}$, is modeled as:

$$
N_{i,t+1} = (N_{i,t}+\phi_i R_t)~exp\left[r_i\left(1-\frac{\alpha_{i1} R_t + \sum_{j=1}^S \alpha_{ij}N_{j,t}}{K_i}\right)\right]~exp(\epsilon_{i,t})
$$

where $r_i$ is the intrinsic growth rate, $\alpha_{ij}$ the competition coefficient of species $j$ on species $i$, $K_i$ the carrying capacity, $R_t$ the number of released individuals for the enhancement of species 1, and $\epsilon_{i,t}$ the species response to stochastic environmental fluctuations that follow a normal distribution $Normal(0,\sigma_{\epsilon}^2)$. To capture variation in species traits, intrinsic growth rates of unenhanced species and interspecific competition coefficients were drawn randomly from a uniform ($r_{i,i \ne 1} \sim Unif(0.5,2)$) and an exponential distribution ($\alpha_{ij,i \ne j} \sim Exp(1/\bar{\alpha})$). The parameter $\phi_i$ controls the relative fitness of captive-bred individuals as follows:

>>>>

Human demands for natural resources are ever-increasing, such that active interventions are critical to the sustainable management of fisheries, forestry, and wildlife [@laikreCompromisingGeneticDiversity2010]. Releases of captive-bred native species ("intentional release") is a form of the efforts to enhance wild populations of diverse taxa [@snyderLimitationsCaptiveBreeding1996; @laikreCompromisingGeneticDiversity2010]. The magnitude of intentional release for commercial and recreational species is massive; typically, a single program releases tens of thousands to millions of plant, invertebrate (bivalves @haagNorthAmericanFreshwater2012, insects), and vertebrate individuals (fish, birds) into the wild annually (reviewed in [@laikreCompromisingGeneticDiversity2010; @kitadaEconomicEcologicalGenetic2018]). There is growing awareness that intentional release entails ecological risks, including the accumulation of deleterious alleles [@arakiGeneticEffectsCaptive2007], the spread of diseases and parasites in wild populations [@krkosekDecliningWildSalmon2007], and the intensification of intraspecific competition within the enhanced species. Yet, this method is still pervasive in natural resource management likely because of the apparent economic benefit of release programs. For example, Kitada @kitadaEconomicEcologicalGenetic2018 analyzed economic performance of release programs for 12 major fishery resources -- the gross monetary yield exceeded the per-seed cost in nine species, suggesting that release programs are economically profitable in general.

Current debates, however, overlook the fact that we have rarely assessed the community-wide impact of intentional release. Species are all embedded in the complex web of interacting organisms, and the stable coexistence of competing species through density-dependent feedback underpins the emergent stability of ecological communities [@tilmanEcologicalConsequencesChanges1999; @carrollNicheFitnessDifferences2011a; @loreauDoesFunctionalRedundancy2004]. Intentional release may disrupt the sensitive balance of species interactions because it introduces unnaturally high numbers of individuals into the wild [@careyNativeInvadersChallenges2012]. Hence, this form of species management may intervene in the ecological process that allows competing species to coexist, ultimately degrading the long-term community stability. In particular, the community-wide impact must be serious in release programs for fishery resouces as many of them release millions of hatchery individuals into a single water body (e.g., a watershed) every year, which is likely far beyond the environmental capacity of the recipient ecosystem. Evidence for this hypothesis is lacking, however.

@@

We employed a multispecies Ricker model [@fowlerSpeciesDynamicsAlter2012] to simulate community dynamics with the intentional release of a constituent species (species 1). Specifically, the population density of species $i$ at time $t+1$, $N_{i,t+1}$, is modeled as:

$$
N_{i,t+1} = (N_{i,t}+\phi_i R_t)\exp\left[r_i\left(1-\frac{\alpha_{i1} R_t + \sum_{j=1}^S \alpha_{ij}N_{j,t}}{K_i}\right)\right]\exp(\varepsilon_{i,t})
$$

where $r_i$ is the intrinsic growth rate, $\alpha_{ij}$ the competition coefficient of species $j$ on species $i$, $K_i$ the carrying capacity, $R_t$ the number of released individuals for the enhancement of species 1, and $\varepsilon_{i,t}$ the species response to stochastic environmental fluctuations that follow a normal distribution $\text{Normal}(0,\sigma_{\varepsilon}^2)$. To capture variation in species traits, intrinsic growth rates of unenhanced species and interspecific competition coefficients were drawn randomly from a uniform ($r_{i,i \ne 1} \sim \text{Unif}(0.5,2.5)$) and an exponential distribution ($\alpha_{ij,i \ne j} \sim \text{Exp}(1/\bar{\alpha})$). The parameter $\phi_i$ controls the relative fitness of captive-bred individuals as follows:
